Syntax and Semantics of Abstract Binding Trees

### 2015-11-11 paper.tex
Use Sigma subscript and update PDF
Don't use elem in fib; looks less trivial this way
[paper] change fiber definition back
[paper] Use a symbol to make abs easier to read
[paper] further clarify fiber/operator stuff

% fonts
\usepackage{pxfonts}
\usepackage{eulervm}
\usepackage[
  activate={true,nocompatibility},
  kerning=true,
  spacing=true,
  tracking=true
]{microtype}
\microtypecontext{spacing=nonfrench}
% \setmainfont{Linux Libertine O}
% \setmathfont{Asana Math}

@@

\newcommand\MkValence[3]{%
  \ifthenelse{\equal{#1}{}}{}{%
    \{#1\}%
  }%
  \ifthenelse{\equal{#2}{}}{}{%
    [#2]%
  }.\, #3%
}
\newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
\newcommand\MkArity[2]{(#1)\, #2}
\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}

@@

\newcommand\IsOperator[3]{
  \IMode{#1}\,{\Vdash}_{\Sigma}\,\IMode{#2}: \OMode{#3}
}

@@

\newcommand\IsAbt[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \,{\IMode{#3}\vdash}_{\scriptscriptstyle{\Operators}}\,%
  \IMode{#4}:\OMode{#5}%
}
\newcommand\IsWf[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}\in_{\mathbf{wf}}\IMode{#5}%
}

\newcommand\IsBTm[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \,{\IMode{#3}\vdash}_{\scriptscriptstyle{\Operators}}\,%
  \IMode{#4}:\IMode{#5}%
}

@@

By the Grothendieck construction
$\Iso{\Grothendieck{}{\Operators}}{\Comma{*}{\Operators}}$ we have a category
of elements of $\Operators$ where objects are triples
$\Pair{\Pair{\Upsilon}{a}}{\vartheta}$, interpreted as
$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$, and morphisms are diagrams:
\[
  \begin{tikzcd}[ampersand replacement = \&]
    \& *
      \arrow[dl, "\vartheta", swap]
      \arrow[dr, "\vartheta'"]
    \&
    \\
    {\Operators\Pair{\Upsilon}{a}}
      \arrow[rr, "\Operators\Pair{\rho}{f}", swap]
    \&
    \& {\Operators\Pair{\Upsilon'}{a'}}
    \\
    {\Pair{\Upsilon}{a}}
      \arrow[rr, "\Pair{\rho}{f}", swap]
    \&
    \& {\Pair{\Upsilon'}{a'}}
  \end{tikzcd}
\]
We will write $\IsOperator{\Upsilon}{\vartheta}{a}$ in case
$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and note that this judgment
enjoys the structural properties of weakening and exchange via functoriality of
$\Operators$. Operators $\vartheta_{0}, \ldots, \vartheta_{n}$ are defined by
specifying the fibers of
$\Of{\pi_{\Operators}}{\Grothendieck{}{\Operators}\to\SCtx\times\Discrete{\Arities}}$
in which they reside:
\begin{alignat*}{3}
  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,\mathrel{\ {{:}}}\,&&\Sets\\
  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,=\,\ && %
    \{\,\vartheta \in \Pair{\Upsilon}{a}%
      {{|}}\ {\pi_{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}} = \Pair{\Upsilon}{a}%
  \}
\end{alignat*}

@@

For instance, consider the lambda calculus with a single sort, $\SortExp$; we
give its signature by asserting the following about its operators:
\begin{gather*}
  \IsOperator{\Upsilon}{\lambda}{
    \MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}
  }\\
  \IsOperator{\Upsilon}{\mathtt{ap}}{
    \MkArity{
      \MkValence{}{}{\SortExp},
      \MkValence{}{}{\SortExp}
    }{\SortExp}
  }
\end{gather*}
These rules correspond to objects of $\Grothendieck{}{\Operators}$ and we
collect their fibers into sets of operators:
\begin{alignat*}{5}
{\pi_{\Operators}}^{-1}&\Pair{\Upsilon}{{\MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}}}\quad&&=
  &&\quad\MkSet{\lambda}\\
{\pi_{\Operators}}^{-1}&\Pair{\Upsilon}{\MkArity{\MkValence{}{}{\SortExp},\MkValence{}{}{\SortExp}}{\SortExp}}\quad&&=
  &&\quad\MkSet{\mathtt{ap}}\\
\bigcup_{\Pair{\Upsilon}{a}}&{{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}}\quad&&=
  &&\quad\MkSet{\lambda,\ \mathtt{ap}}\\
\end{alignat*}
So far, we have made no use of symbols and parameters; however, consider the
extension of the calculus with assignables (references):

>>>>

% fonts
\usepackage{pxfonts}
\usepackage{eulervm}
\usepackage{bbold}
\usepackage[
  activate={true,nocompatibility},
  kerning=true,
  spacing=true,
  tracking=true
]{microtype}
\microtypecontext{spacing=nonfrench}
% \setmainfont{Linux Libertine O}
% \setmathfont{Asana Math}

@@

\newcommand\MkValence[3]{%
  \ifthenelse{\equal{#1}{}}{}{%
    \{#1\}%
  }%
  \ifthenelse{\equal{#2}{}}{}{%
    [#2]%
  }.\, #3%
}
\newcommand\MkBTm[3]{\mathbb{\lambda}\{#1\}[#2].\, #3}
\newcommand\MkArity[2]{(#1)\, #2}
\newcommand\IsArity[1]{\IMode{#1}\ \textit{arity}}
\newcommand\IsValence[1]{\IMode{#1}\ \textit{valence}}
\newcommand\IsSort[1]{\IMode{#1}\ \textit{sort}}
\newcommand\Leq[2]{\IMode{#1}\leq\IMode{#2}}

@@

\newcommand\IsOperator[3]{
  \IMode{#1}\,{\Vdash}\,\IMode{#2}: \OMode{#3}
}

@@

\newcommand\IsAbt[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \,{\IMode{#3}\vdash}\,%
  \IMode{#4}:\OMode{#5}%
}
\newcommand\IsWf[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \IMode{#3}\vdash%
  \IMode{#4}\in_{\mathbf{wf}}\IMode{#5}%
}

\newcommand\IsBTm[5]{%
  \IMode{#1}\triangleright%
  \IMode{#2}\parallel%
  \,{\IMode{#3}\vdash}\,%
  \IMode{#4}:\IMode{#5}%
}

@@

By the Grothendieck construction
$\Iso{\Grothendieck{}{\Operators}}{\Comma{*}{\Operators}}$ we have a category
of elements of $\Operators$ where objects are triples
$\Pair{\Pair{\Upsilon}{a}}{\vartheta}$, interpreted as
$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$, and morphisms are diagrams:
\[
  \begin{tikzcd}[ampersand replacement = \&]
    \& *
      \arrow[dl, "\vartheta", swap]
      \arrow[dr, "\vartheta'"]
    \&
    \\
    {\Operators\Pair{\Upsilon}{a}}
      \arrow[rr, "\Operators\Pair{\rho}{f}", swap]
    \&
    \& {\Operators\Pair{\Upsilon'}{a'}}
    \\
    {\Pair{\Upsilon}{a}}
      \arrow[rr, "\Pair{\rho}{f}", swap]
    \&
    \& {\Pair{\Upsilon'}{a'}}
  \end{tikzcd}
\]
We will write $\IsOperator{\Upsilon}{\vartheta}{a}$ in case
$\Member{\vartheta}{\Operators\Pair{\Upsilon}{a}}$ and note that this judgment
enjoys the structural properties of weakening and exchange via functoriality of
$\Operators$. Operators $\vartheta_{0}, \ldots, \vartheta_{n}$ are defined by
specifying the fibers of
$\Of{\pi_{\Operators}}{\Grothendieck{}{\Operators}\to\SCtx\times\Discrete{\Arities}}$
in which they reside:
\begin{alignat*}{3}
  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,\mathrel{\ {{:}}}\,&&\SCtx\times\Discrete{\Arities}\to\Sets\\
  &{\pi_{\Operators}}^{-1}{\Pair{\Upsilon}{a}}&&\,=\,\ && %
    \{\ \ \Member{\vartheta}{\Pair{\Upsilon}{a}}\ \ |\ \
    {\pi_{\Operators}}{\Pair{\Pair{\Upsilon}{a}}{\vartheta}} =
    \Pair{\Upsilon}{a}\,\}
\end{alignat*}

@@

For instance, consider the $\lambda$-calculus with a single sort, $\SortExp$; we
give its signature $\Sigma_{\lambda}$ by asserting the following about its
operators:
\begin{gather*}
  \IsOperator{\Upsilon}{\mathtt{lam}}{
    \MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}
  }\\
  \IsOperator{\Upsilon}{\mathtt{ap}}{
    \MkArity{
      \MkValence{}{}{\SortExp},
      \MkValence{}{}{\SortExp}
    }{\SortExp}
  }
\end{gather*}
These rules correspond to objects of $\Grothendieck{}{{\Operators}_{\lambda}}$ and we
collect their fibers into sets of operators:
\begin{alignat*}{5}
{\pi_{{\Operators}_{\lambda}}}^{-1}&\Pair{\Upsilon}{{\MkArity{\MkValence{}{\SortExp}{\SortExp}}{\SortExp}}}\quad&&=
  &&\quad\MkSet{\mathtt{lam}}\\
{\pi_{{\Operators}_{\lambda}}}^{-1}&\Pair{\Upsilon}{\MkArity{\MkValence{}{}{\SortExp},\MkValence{}{}{\SortExp}}{\SortExp}}\quad&&=
  &&\quad\MkSet{\mathtt{ap}}\\
{\pi_{{\Sigma}_{\lambda}}}^{-1}&=\bigcup_{\Pair{\Upsilon}{a}}{{\pi_{{\Operators}_{\lambda}}}^{-1}{\Pair{\Upsilon}{a}}}\quad&&=
  &&\quad\MkSet{\mathtt{lam},\ \mathtt{ap}}\\
\end{alignat*}
So far, we have made no use of symbols and parameters; however, consider the
extension of the calculus with assignables (references):
